A Generalization of Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-03-26 reports/paper-rda.Rnw
Finished first draft of paper.

\begin{abstract}
\cite{Friedman:1989tm} has proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting. This \emph{RDA} classifier employs a regularized covariance-matrix estimator, which incorporates a ridge-like shrinkage factor by adding a positive constant to the zero-eigenvalue estimates to ensure positive definiteness. Here, we propose a novel generalization of Friedman's \emph{RDA} classifier that is more computationally efficient than its original counterpart because the inverses and determinants computed in the decision rule can be reduced from the original feature space to a linear subspace with dimensions corresponding to the individual class sample sizes. Additionally, our generalization of the \emph{RDA} (\emph{GRDA}) classifier enables efficient estimation of optimal tuning parameters from a large set of candidate values, simplifies the construction of the regularized covariance-matrix estimator to improve its interpretation, and allows for a more flexible classification model. Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers. Finally, we provide an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R package, available on CRAN.
\end{abstract}

@@

Microarrays have enabled researchers to measure gene-expression levels of thousands of genes in a single experiment to identify a subset of genes and their expression patterns that either characterize a particular cell state or facilitate an accurate prediction of future cell state.  We focus on the classification of tissues and other biological samples into one of $K$ disjoint, predefined cancer types, where our goal is to ascertain the class membership of a new tissue sample based on gene-expression microarray data. However, microarray data typically consist of a large number of gene expressions $p$ relative to the sample size $N$. Moreover, standard benchmark microarray data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer than $N = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance, due to the \emph{curse of dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample covariance matrices employed in both classifiers are singular. As a result, many researchers, including \cite*{Mai:2012bf}, \cite*{Merchante:2012vk}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, \cite*{Huang:2010ju}, and \cite*{Clemmensen:2011kr}, have introduced restrictions on the population covariance matrices in the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.  For instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA} (\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.

@@

Here, we consider a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed a weighted average of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers. \cite{Friedman:1989tm} has additionally incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA} classifier is well-posed. \cite{Friedman:1989tm} has shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $p > N$. However, in its original form, the \emph{RDA} classifier is impractical when $p \gg N$ because the \emph{RDA} decision rule requires the calculation of the inverses and determinants of the $p \times p$ covariance-matrix estimators for each class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are selected via cross-validation, which adds a substantial computational burden to the model-selection process. The \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires $O(Kp^3)$ calculations for each tuning-parameter pair for each fold in the cross-validation scheme.

In this paper we address the computational shortcomings of the \emph{RDA} classifier. Applying a technique similar to \cite{Ye:2006tq}, we decompose the \emph{RDA} classifier and show that the sum related to the null space of the pooled sample covariance matrix can be discarded. In turn, this omission provides a substantial increase in the computational efficiency of the \emph{RDA} classifier and its model-selection. In fact, we show rigorously that the inverse and determinants can be reduced to calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$. This reduction in calculation is clearly advantageous because $n_k$ is often quite small, say, 15--20, in microarray data sets. Thus, we demonstrate that the calculation of the \emph{RDA} decision rule involving inverses and determinants of extremely large covariance matrices is equivalent a decision rule involving inverses and determinants on much smaller matrices.

Also, for our \emph{GRDA} classifier we first simplify the pooling matrix in the \emph{RDA} classifier by constructing instead a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers. This simplification is advantageous in that it improves the interpretation of the pooling operation of the \emph{RDA} classifier. We show that under mild regularity conditions the \emph{GRDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix. Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues. Although \cite{Ji:2008wp} and \cite{Ye:2006jm} have provided a similar invariance property for the \emph{LDA} decision rule, the \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent in practice \citep{Clemmensen:2011kr}. The \emph{GRDA} classifier allows us to relax this assumption by including unrestricted covariance matrices in the model. The \emph{GRDA} classifier is then a compromise between the two approaches, favoring the stronger linear assumption if a larger value of the pooling parameter is selected as optimal. Our formulation provides a general framework from which we can elucidate properties of covariance-regularization methods and the relationships among them in the linear case, the quadratic case, and the cases in between.

Many of the recently proposed classifiers have a variable-selection method integrated. While this is advantageous in potentially determining biomarkers that are correlated with the classification label, these methods often yield poor classification accuracy if a large number of variables are retained. Moreover, these methods can require large computational costs, making these methods undesirable in practice. To study these issues in depth, we consider several simulated data sets, where we show that the classification degrades as the number of features is increased. Contrarily, our proposed \emph{GRDA} classifier handles the large number of features well because our included dimension-reduction method retains only the classificatory information in the covariance-matrix estimators. Moreover, while the competing classifiers exhibit large computational costs as the number of features in the original data set is increased, the \emph{GRDA} classifier handles the larger number of features efficiently.

In addition to the improvement in computational performance of classifiers with regularized covariance matrices, we demonstrate that our \emph{GRDA} classifier is competitive with and often superior to several recent classifiers designed for small-sample, high-dimensional data. In particular, we compare our \emph{GRDA} classifier with an alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of the \emph{DLDA} classifier from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}. We conduct our comparison of the classification accuracy of the competing classifiers using four small-sample, high-dimensional microarray data sets.

We have organized the remainder of the paper as follows. In Section 2 we present our generalization of the \emph{RDA} classifier after reviewing its original counterpart from \cite{Friedman:1989tm}. In Section 3 we show that our proposed classifier includes several well-known regularized classifiers as special cases and briefly discuss some commonalities among them. In Section 4 we describe four small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy. We then conclude with a brief discussion in Section 5.

@@

We also require the following additional notation throughout the document. Let $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v = 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$. Let $\bm I_m$ be the $m \times m$ identity matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$.

@@

As we have discussed, several eigenvalue adjustment methods have been proposed to increase eigenvalues (approximately) equal to 0. Here, we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \bm \Phi_k,\label{eq:grda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\bm \Phi_k \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix consisting of the eigenvalue shrinkage constants. Thus, we have the \emph{GRDA} (\emph{GRDA}) classifier
\begin{align}
	D_{GRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|, \label{eq:grda}
\end{align}
The specification of $\alpha_k$ and $\bm \Phi_k$ in \eqref{eq:grda-cov} allows for a large family of eigenvalue-shrinkage methods and their relationships to be examined in depth.

@@

In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of $\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of $\widehat{\bm \Sigma}$,  the columns of $\bm U \in \mathbb{R}_{p \times p}$ are the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$. Then, we partition $\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$. We also partition $\bm \Phi_k = \bm \Phi_{k1} \oplus \bm \Phi_{k2}$ such that $\bm \Phi_{k1} \in \mathbb{R}_{q \times q}$ and $\bm \Phi_{k2} \in \mathbb{R}_{(p-q) \times (p-q)}$.

\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) = q \le p$. Then, we have
\begin{align}
	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \bm \Phi_{k2})\bm U',\label{eq:rda-matrix}
\intertext{where}
\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \bm \Phi_{k1}.\label{eq:Wk}
\end{align}
\end{lemma}
\begin{proof}
From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 = \bm 0_{p \times (p - q)}$. Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$. Then, notice that $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \bm U' \bm \Phi_k \bm U$. Thus, \eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
\end{proof}

@@

\begin{cor}
Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in \eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$, rank$(\widehat{\bm \Sigma}_k(\lambda)) = q$ for all $k = 1, \ldots, K$.
\end{cor}
\begin{proof}
The proof follows by setting $\bm \Phi_k = \bm 0_p$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
\end{proof}

@@

Now, we present our main result, where we decompose the \emph{GRDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification performed using Lemma \ref{lemma:RDA-constant-term}. To reduce the computational costs of the \emph{GRDA} classifier, we require that $\bm \Phi_{k2} = \bm \Phi_2$, $k = 1, \ldots, K$.
Here, we present our main result that reduces \eqref{eq:grda} to an equivalent, more computationally efficient decision rule.

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix} and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Additionally, let $\bm \Phi_{k2} = \bm \Phi_2$, $k = 1, \ldots, K$. Then, the \emph{GRDA} decision rule in \eqref{eq:grda} is equivalent to the reduced form
	\begin{align}
		D_{GRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:rda-decomposed}
	\end{align}

\end{thm}
\begin{proof}
	From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{-1} = \bm U(\bm W_k^{-1} \oplus\bm \Phi_{k2}^{-1})\bm U'$ and $|\tilde{\bm \Sigma}_k| = | \bm W_k | |\bm \Phi_{k2}|$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
	\begin{align*}
	(\bm x - \xbar_k)' \tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k| &= (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)\\
	&+ (\bm x - \xbar_k)' \bm U_2 \bm \Phi_{k2}^{-1} \bm U_2' (\bm x - \xbar_k) + \ln | \bm W_k | + \ln |\bm \Phi_{k2}|.
	\end{align*}
From Lemma \ref{lemma:RDA-constant-term}, $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$. Therefore, the proof follows because $\bm \Phi_{k2} = \bm \Phi_2$.
\end{proof}

@@

\begin{proposition}\label{proposition:rda-W_k}
Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$ and
	\begin{align}
		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
	\end{align}
where $\bm Q_k = \bm I_{n_k} + \alpha_k(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q + \bm \Phi_{k1}$.
\end{proposition}
\begin{proof}
	To calculate $|\bm W_k|$, from \cite{Harville:2008wja} we apply Theorem 18.1.1, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Also,  \eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula \citep[Theorem 18.2.8]{Harville:2008wja}.
\end{proof}

@@

@@

Approximately 30-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of DLBCL patients are cured with modern chemotherapeutic regimens.  Several models have been proposed, such as the International Prognostic Index (IPI), to determine a patient's curability.  These models rely on clinical covariates, such as age, to determine if the patient can be cured and are often ineffective. \cite{Shipp:2002ka} have argued that more effective means are desired to determine a patient's curability. The authors measured 6,817 gene expression levels from 58 DLBCL patient samples with customized cDNA (‘lymphochip’) microarrays to investigate the curability of patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based chemotherapy.  Among the 58 DLBCL patient samples, 32 are from cured patients while 26 are from patients with fatal or refractory disease.

>>>>

\begin{abstract}

A reliable classification of tumors from gene expression microarray data is essential for successful diagnosis and treatment of diseases such as cancer, but the analysis of microarray data often presents difficult challenges because the number of gene expression levels is typically much larger than the sample size. Consequently, several sparse and regularized classifiers have been proposed often with strict covariance assumptions or eigenvalue adjustments to ensure that the classifiers are well-posed. From \cite{Friedman:1989tm} we consider specifically the popular regularized discriminant analysis (\emph{RDA}) classifier, which requires computationally expensive matrix operations and involves a time-consuming model-selection procedure, making the \emph{RDA} classifier impractical for high-dimensional data. We propose a novel generalization of the \emph{RDA} (\emph{GRDA}) classifier that improves upon the interpretation of the \emph{RDA} classifier and incorporates a more general covariance-matrix regularization approach, which includes as special cases several proposed sparse and regularization methods. Furthermore, we address the computational shortcomings of the \emph{RDA} classifier by integrating a novel dimension-reduction technique that enables a computationally efficient model-selection procedure. Moreover, we demonstrate with several high-dimensional microarray data sets that our proposed classifier is competitive and often superior to several recently proposed sparse and regularized classifiers in terms of classification accuracy. Finally, we provide an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R package, available on CRAN.

\end{abstract}

@@

Microarrays have enabled researchers to measure gene-expression levels of thousands of genes in a single experiment to identify a subset of genes and their expression patterns that either characterize a particular cell state or facilitate an accurate prediction of future cell state.  We focus on the classification of tissues and other biological samples into one of $K$ disjoint, predefined cancer types, where our goal is to ascertain the class membership of a new tissue sample based on gene-expression microarray data. However, microarray data typically consist of a large number of gene expressions $p$ relative to the sample size $N$. Moreover, standard benchmark microarray data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer than $N = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance, due to the \emph{curse of dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample covariance matrices employed in both classifiers are singular. As a result, many researchers, including \cite*{Mai:2012bf}, \cite*{Merchante:2012vk}, \cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju}, have introduced restrictions on the population covariance matrices in the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.  For instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA} (\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.

@@

Here, we consider a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers. \cite{Friedman:1989tm} has additionally incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA} classifier is well-posed. \cite{Friedman:1989tm} has shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $p > N$. However, in its original form, the \emph{RDA} classifier is impractical when $p \gg N$ because the \emph{RDA} decision rule requires the calculation of the inverses and determinants of the $p \times p$ covariance-matrix estimators for each class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are selected via cross-validation, which adds a substantial computational burden to the model-selection process. The \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires $O(Kp^3)$ calculations for each tuning-parameter pair.

In this paper we provide a generalization of the \emph{RDA} classifier by first simplifying the pooling matrix employed in the \emph{RDA} classifier. We construct instead a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers. We demonstrate that this simplification improves the interpretation of the pooling operation of the \emph{RDA} classifier. The \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent in practice \citep{Clemmensen:2011kr}. The \emph{GRDA} classifier allows us to relax the linearity assumption by including unequal covariance matrices in the model. The \emph{GRDA} classifier is then a compromise between the \emph{LDA} and \emph{QDA} classifiers,  so that the \emph{GRDA} classifier includes both the \emph{LDA} and \emph{QDA} classifiers as special cases. We then introduce a more general shrinkage term to the \emph{RDA} classifier to establish our \emph{GRDA} classifier. Our formulation of covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2013ji} and \cite*{Xu:2009fl}.

We show that the \emph{GRDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix. Applying a technique similar to \cite{Ye:2006tq}, we decompose the \emph{RDA} classifier and show that the sum related to the null space of the pooled sample covariance matrix can be discarded. Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues. This invariance property addresses the computational shortcomings of the \emph{RDA} classifier and provides a substantial increase in the computational efficiency of the \emph{RDA} classifier and its model selection. In fact, we show that the inverse and determinants can be reduced to calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$. This reduction in calculation is clearly advantageous because $n_k$ is often quite small, say, 15--20, in microarray data sets. Thus, through a dimension-reduction technique we demonstrate that the calculation of the \emph{RDA} decision rule involving inverses and determinants of extremely large covariance matrices is equivalent a decision rule involving inverses and determinants on much smaller matrices.

In addition to the improvement in computational performance, we demonstrate that our \emph{GRDA} classifier is competitive with and often superior to several recent classifiers designed for small-sample, high-dimensional data. In particular, we compare our \emph{GRDA} classifier with the \emph{DLDA} classifier from \cite{Dudoit:2002ev} and the recently proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and \cite{Witten:2011kc}. We conduct our comparison of the classification accuracy of the competing classifiers using four small-sample, high-dimensional microarray data sets.

We have organized the remainder of the paper as follows. In Section 2 we review the \emph{RDA} classifier from \cite{Friedman:1989tm} and discuss our generalization in Section 3. In Section 4 we describe four small-sample, high-dimensional microarray data sets and then demonstrate that our proposed \emph{GRDA} classifier is often superior to the competing classifiers in terms of classification accuracy. We then conclude with a brief discussion in Section 5.

@@

We also require the following additional notation throughout the document. Let $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v = 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$. Let $\bm I_m$ be the $m \times m$ identity matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$. Lastly, for $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.

@@

As we have discussed, several eigenvalue adjustment methods have been proposed to increase eigenvalues (approximately) equal to 0. Here, we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:grda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage constant. Thus, we have the \emph{GRDA} classifier
\begin{align}
	D_{GRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|. \label{eq:grda}
\end{align}
For $\gamma > 0$, $\tilde{\bm\Sigma}_k$ is nonsingular such that $\tilde{\bm\Sigma}_k^{-1}$ can be substituted for $\tilde{\bm\Sigma}_k^{+}$ in \eqref{eq:grda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm\Sigma}_k|$ equal to the product of the positive eigenvalues of $\tilde{\bm\Sigma}_k$.

@@

In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of $\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p \times p}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$ with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of $\widehat{\bm \Sigma}$,  the columns of $\bm U \in \mathbb{R}_{p \times p}$ are the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and rank$(\widehat{\bm \Sigma}) = q$. Then, we partition $\bm U = (\bm U_1, \bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in \mathbb{R}_{p \times (p - q)}$.

\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U'$ be the eigendecomposition of $\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) = q \le p$. Then, we have
\begin{align}
	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U',\label{eq:rda-matrix}
\intertext{where}
\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_{q}.\label{eq:Wk}
\end{align}
\end{lemma}
\begin{proof}
From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space of $\widehat{\bm \Sigma}$, which implies that $\widehat{\bm \Sigma}_k \bm U_2 = \bm 0_{p \times (p - q)}$. Hence, for $k = 1, \ldots, K$, $\bm U' \widehat{\bm \Sigma}_k \bm U = \bm U_1' \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$. Then, notice that $\bm U' \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U' \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$. Thus, \eqref{eq:rda-matrix} follows because $\bm U$ is orthogonal.
\end{proof}

@@

\begin{cor}
Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in \eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$, rank$(\widehat{\bm \Sigma}_k(\lambda)) = q$ for all $k = 1, \ldots, K$.
\end{cor}
\begin{proof}
The proof follows by setting $\gamma = 0$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
\end{proof}

@@

Now, we present our main result, where we decompose the \emph{GRDA} decision rule and show that the component requiring the largest computational costs does not contribute to the classification performed using Lemma \ref{lemma:RDA-constant-term}. Here, we present our main result that reduces \eqref{eq:grda} to an equivalent, more computationally efficient decision rule.

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix} and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then, the \emph{GRDA} decision rule in \eqref{eq:grda} is equivalent to the reduced form
	\begin{align}
		D_{GRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:rda-decomposed}
	\end{align}

\end{thm}
\begin{proof}
	From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(\bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U'$ and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q}| \bm W_k |$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
	\begin{align*}
	(\bm x - \xbar_k)' \tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k| &= (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k)\\
	&+ \gamma^{+} (\bm x - \xbar_k)' \bm U_2 \bm U_2' (\bm x - \xbar_k) + \ln | \bm W_k | + (p - q) \ln \gamma.
	\end{align*}
Therefore, the proof follows from Lemma \ref{lemma:RDA-constant-term} because $\bm U_2' (\bm x - \xbar_k)$ is constant for all $k = 1, \ldots, K$.
\end{proof}

@@

\begin{proposition}\label{proposition:rda-W_k}
Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$ and
	\begin{align}
		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1' \bm X_k' \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
	\end{align}
where $\bm Q_k = \bm I_{n_k} + \alpha_k(1- \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1' \bm X_k'$ and $\bm \Gamma_k = \alpha_k \lambda \bm D_q + \gamma \bm I_q$.
\end{proposition}
\begin{proof}
	To calculate $|\bm W_k|$, from \cite{Harville:2008wja} we apply Theorem 18.1.1, which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm U_1' \bm X_k'$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Also, \eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula \citep[Theorem 18.2.8]{Harville:2008wja}.
\end{proof}

@@

\[ \left[\text{Table 1 goes about here }\right]\]

@@

Approximately 30-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of DLBCL patients are cured with modern chemotherapeutic regimens.  Several models have been proposed, such as the International Prognostic Index (IPI), to determine a patient's curability.  These models rely on clinical covariates, such as age, to determine if the patient can be cured and are often ineffective. \cite{Shipp:2002ka} have argued that more effective means are desired to determine a patient's curability. The authors measured 6,817 gene expression levels from 58 DLBCL patient samples with customized cDNA (lymphochip) microarrays to investigate the curability of patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based chemotherapy.  Among the 58 DLBCL patient samples, 32 are from cured patients while 26 are from patients with fatal or refractory disease.
